### 2016-06-14 capability-machine.tex
Added simple intrastructure for comments. Added specification of start of proof.

\declaretheorem[numbered=yes,name=Lemma]{lemma}
\declaretheorem[numbered=yes,name=Definition]{definition}

@@

@@

%LR
\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\envintr}[1]{\intr{G}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdgr}{\envintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\observations}{\mathcal{O}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}

@@

Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$. Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).

@@

@@

    \init: & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\

@@

           & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\

@@

Here $\size$ is 9. The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities. $i_1 \dots i_7$ refers to the instructions in the following program:

@@

\begin{lemma}
 Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.

@@

@@

Register-file relation:
\begin{align*}
  \stdgr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
  \stdgr(W) \defeq & \{ \npair{\reg} \mid \\
                    & \quad \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
\end{align*}

``Continuation'' relation:
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stdkr(W) \defeq & \{ \npair{c} \mid \\
                   & \quad \forall v \in \stdvr(W)\ldotp \\
                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat[\heap]{n'}{W'} \ldotp \forall \reg \in \stdgr(W') \\
                   & \qquad \quad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
\end{align*}
``Expression'' relation:
\begin{align*}
  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
  \stder(W) \defeq & \{ \npair{\pc} \mid \\
                   & \quad \forall \reg \in \stdgr(W) \ldotp \\
                   & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\
                   & \qquad \quad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
                   & \qquad \qquad (n,\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f) \in \observations(W) \}
\end{align*}

>>>>

\declaretheorem[numbered=yes,name=Lemma]{lemma}
\declaretheorem[numbered=yes,name=Definition]{definition}
\declaretheorem[numbered=yes,name=Specification]{specification}

@@

%Comments
\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}}
\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}}

@@

%LR
\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
\newcommand{\valueintr}[1]{\intr{V}{#1}}
\newcommand{\exprintr}[1]{\intr{E}{#1}}
\newcommand{\contintr}[1]{\intr{K}{#1}}
\newcommand{\regintr}[1]{\intr{R}{#1}}
\newcommand{\stdvr}{\valueintr{\asmType}}
\newcommand{\stder}{\exprintr{\asmType}}
\newcommand{\stdrr}{\regintr{\asmType}}
\newcommand{\stdkr}{\contintr{\asmType}}
\newcommand{\observations}{\mathcal{O}}
\newcommand{\npair}[2][n]{\left(#1,#2 \right)}

@@

Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$. Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0). The following is a test program for a ticket dispenser:

@@

\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\

@@

     & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\

@@

@@

Here $\size$ is 9. The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities. $i_1 \dots i_7$ refers to the instructions in the following ticket dispenser:

@@

\begin{lemma}[Ticket dispenser test]
\label{lem:tckt-disp}
 Given $\Phi$, if $\memheap$ contains the ticket dispenser test program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.

@@

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it obeys the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc]
  For all worlds, $W$, configurations, $\Phi$, natural numbers, $n$, and addresses $b$, $e$, and $return$ where $\heapSat[\memheap]{n}{W}$ if
  \begin{itemize}
  \item $\memreg(r_0) = (\entry,b,e,\var{return})$,
  \item $\memreg(r_1) = m$ where $m$ is some positive natural number\footnote{We do not include 0 as malloc in C returns a null pointer when malloc(0) is called.}, and
  \item the remaining registers contain 0,
  \end{itemize}
then there exists a configuration $\Phi'$, a future world $W' \future W$, a natural number $i$ such that $\heapSat[{\memheap[\Phi']}]{n-i}{W'}$ and $\dom(\memheap[\Phi']) \supseteq \dom(\memheap) \uplus [b',e']$\footnote{notice that $\malloc$ is allowed to allocated memory for its own purposes} for some addresses $b'$ and $e'$. Further the register file contains the following
\begin{itemize}
\item $\memreg(\pcreg) = (\rwx,b,e,\var{return})$,
\item $\memreg(r_0) = (\entry,b,e,\var{return})$,
\item $\memreg(r_1) = (\rwx,b',e',b')$, and
\item $\memreg(r) = 0$ for all other registers $r$.
\end{itemize}
\end{specification}
\begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
Let configuration $\Phi$ be given. Assume that the ticket dispenser test code resides in memory starting at address $\init$. Further assume that the register file contains the following:
\begin{enumerate}
\item $\memreg(\rio) = (\readwrite,0,0,0)$ 
\item $\memreg(\pcreg) = (\rwx,\init, \init + \var{proglen} - 1, \init)$
\item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
\end{enumerate}
Further assume $\Phi \rightarrow^* (\halted,\Phi')$.
\end{proof}

@@

Register-file relation:
\begin{align*}
  \stdrr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
  \stdrr(W) \defeq & \{ \npair{\reg} \mid \\
                    & \quad \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
\end{align*}

``Continuation'' relation:
\begin{align*}
  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
  \stdkr(W) \defeq & \{ \npair{c} \mid \\
                   & \quad \forall v \in \stdvr(W)\ldotp \\
                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
                     \forall \heapSat[\heap]{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
                   & \qquad \quad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
\end{align*}
``Expression'' relation:
\begin{align*}
  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
  \stder(W) \defeq & \{ \npair{\pc} \mid \\
                   & \quad \forall \reg \in \stdrr(W) \ldotp \\
                   & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\
                   & \qquad \quad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
                   & \qquad \qquad (n,\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f) \in \observations(W) \}
\end{align*}
